---
title: "HW7_MandMs"
---

In this assignment, you'll be picking up where you left off with the M&Ms activity. You'll load the data and perform some basic statistical analysis. Follow the steps below to complete the assignment.

### Task 1: Load the tidyverse
```{r}
library(tidyverse)

```



### Task 2: Load the M&Ms dataset from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/mandms.csv
```{r}
mandm <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/mandms.csv")
```



### Task 3: Sample 5 rows of data from the dataset, and then sample 10 rows, saving each to separate dataframes called sample5 and sample10. Be sure to use set.seed(1234) before you start.

```{r}
sample_5 <- mandm |> 
  sample_n(5)

sample_10 <- mandm |> 
  sample_n(10)

```

### Task 4: Calculate the mean and standard deviation of the two samples, using one of the colors.
```{r}
sample_5 |> summarise(mean = mean(black), sd = sd(black))
sample_10 |> summarise(mean = mean(black), sd = sd(black))
```


### Task 5: Calculate the mean and standard deviation of the original dataset, using the same color as Task 4
```{r}
mandm |> summarise(mean = mean(black), sd = sd(black))
```


### Task 6: Make three histograms: one each for sample5, sample10 and the original dataset. Use a consistent binwidth.
```{r}
mandm |> 
  ggplot() +
  geom_histogram(aes(x = black), binwidth = 3)

sample_5 |> 
  ggplot() +
  geom_histogram(aes(x = black), binwidth = 3)

sample_10 |> 
  ggplot() +
  geom_histogram(aes(x = black), binwidth = 3)

```


### Task 7: Answer some questions

Compare the three sets of means, standard deviations and histograms. Which are more helpful to you in understanding the accuracy of the samples?

I think the original dataset tells us more because it is larger

What does this exercise tell you about the relationship between sample size and estimate accuracy?

Larger datasets are more accurate

How does random sampling introduce variation into statistical estimates?

Random sampling enables a random mix of datapoints and different means/sd

If you repeated this process, would the samples always be identical? Why or why not?

No because the samples are random and have different variations


When finished, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
