Tests I could run:
- mean and standard deviation for tuition, graduation rates, admission rates
- linear correlation
- Z-scores to determine outliers and compare
- scatter plot to visualize trends


```{r}
library(tidyverse)
```


```{r load-data}
college_grad <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/colleges_admit_grads.csv")
```

```{r}
college_grad |> summarise(mean = mean(TUITIONFEE_IN), sd = sd(TUITIONFEE_IN))
```
This is not giving me a value and I am not sure why

```{r}
college_grad |> summarise(mean = mean(TUITIONFEE_OUT), sd = sd(TUITIONFEE_OUT))
```
Same with this and I am not sure what the difference between these tuition fees are

```{r}
college_grad |> summarise(mean = mean(COSTT4_A), sd = sd(COSTT4_A))
```
I believe this set of data is the cost, but I am not sure what the difference of this set and the other 2 tuition sets are. Regardless, the mean is 29,298.41. And the standard deviation is 17,702.59.

```{r}
college_grad |> summarise(mean = mean(ADM_RATE), sd = sd(ADM_RATE))
```
No variable again

```{r}
college_grad |> summarise(mean = mean(ADM_RATE_ALL), sd = sd(ADM_RATE_ALL))
```
??

```{r}
college_grad |> summarize(correlation = cor(COSTT4_A, ADM_RATE, method = "pearson"))
```
Still no variable

```{r}
college_grad |> 
  ggplot() +
  geom_point(aes(x=ADM_RATE, y=COSTT4_A)) +
  geom_smooth(aes(x=ADM_RATE, y=COSTT4_A), method= "lm")
```
I want to replace admission rate with graduation rate but I am having a hard time discerning where that is in the data.


Confused about how to calculate z-scores
