
```{r}
library(tidyverse)
```


```{r load-data}
college_grad <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/colleges_admit_grads.csv")
```


```{r}
college_grad |> summarise(mean = mean(COSTT4_A), sd = sd(COSTT4_A))
```
The mean cost for attending university is $29,298.41. There are some schools that have a much higher cost, but most of the data is clustered around the mean. The standard deviation is pretty high, which proves this as well.

```{r}
college_grad |> filter(!is.na(OMAWDP8_ALL_POOLED_SUPP)) |> summarise(mean = mean(OMAWDP8_ALL_POOLED_SUPP), sd = sd(OMAWDP8_ALL_POOLED_SUPP))
```
The mean gradation rate across all of the universities is 49%. This is pretty low.


```{r}
college_grad |> 
  ggplot() +
  geom_point(aes(x=OMAWDP8_ALL_POOLED_SUPP, y=COSTT4_A)) +
  geom_smooth(aes(x=OMAWDP8_ALL_POOLED_SUPP, y=COSTT4_A), method= "lm")
```
Based on this graph, it does seem like there is reason to believe that higher cost does not impact graduation rate, at least not very much. The data points at the top signify higher costs, and most of them are clustered to the right which indicates higher rate of graduation. But, there is a large number of data points that are much lower, but still towards the right, indicating lower costs but still high rates of graduation. I would guess that the data points in the far top right are more prestigious schools like the ivies, of state schools. I do not believe that there is a high correlation between cost and graduation rate based on this data.

```{r}
college_grad |> summarize(correlation = cor(OMAWDP8_ALL_POOLED_SUPP, COSTT4_A, method = "pearson", use = "complete.obs"))
```
This test determined the correlation between the two variables. It is 66%. This is not very high, proving that the correlation between cost and graduation rate is not significant. 

For my story pitch I would write about how the data suggests that a higher cost/more prestigious school does not necessarily mean that there is a significantly higher rate of graduation. The correlation between the two variable is 66%, further proving this point. The scatterplot also shots many lower cost schools producing high graduation rates. I would look for outliers to help determine which schools are performing the best/worst and how that correlates with my theory. Because of the large number of data points, it may be beneficial to narrow down the data and make it less complex. 
